### 2021-11-07 NSF2021/Description/specificaim2.tex
working on SA2

\subsection{Specific Aim 2: Efficient, high-order methods for
large-scale simulations}
% ----------------------------------------------------------------------
Over the past decade, there has been an explosion of interest in
small-scale processes that utilize capillary forces and van der Waals
interactions to coordinate movement and bind microscopic components in
solvent~\cite{Pandey2011, Zhang2017, Siontorou2017}. Additionally,
hydrodynamic interactions are central to fabricate complex,
three-dimensional microstructures~\cite{Dasgupta2017, Leong2007,
Reynolds2019, Cho2010}. Hydrodynamic effects cannot be ignored since
they, like for the rates of biological functions like pore
dynamics~\cite{RYHAM20112929}, set the rate of dissipation. In Specific
Aim 2, we propose to develop efficient high-order numerical methods to
simulate dynamic assembly of particles under HAP potential in a viscous
solvent. As described in~\eqref{eq:stokes} and illustrated in
Figure~\ref{fig:flow_map}, the solvent phase is modeled by the Stokes
equations for an incompressible, viscous fluid, and these equations are
coupled to the screened-Laplace equation~\eqref{SL} through viscous and
hydrophobic stress balance. 
\todo[inline]{BQ: Also will have the new Allen-Cahn type equation}

\label{subsec:specific_aim_2}
\begin{wrapfigure}[14]{r}{0.61\textwidth}
  \centering{\includegraphics[width=0.60\textwidth]{Figures/domain.pdf}}
    \vspace{-15pt}
    \caption{\label{fig:flow_map} \footnotesize A collection of 3D
    amphiphilic particles suspended in a solvent. The solvent velocity
    satisfies the Stokes equations, and the hydrophobic forces depend on
    the solvent activity that satisfies the screened Laplace equation.
    The geometry is updated by solving the mobility problem.} 
\end{wrapfigure}
The HAP model requires solving exterior Dirichlet problems for the
screened Laplace equation and the Stokes mobility problem at each time
step in complex domains such as in Figure~\ref{fig:domain}. If
discretizing these equations with stencil-based numerical methods, the
computational domain must be truncated, the fluid volume must be
discretized, and artificial boundary conditions must be imposed. The
artificial boundary conditions introduce additional error, and
discretizing the volume results in an excessively large system of
equations. Moreover, it is difficult to obtain high-order
discretizations when the boundary surfaces are irregular. Alternatively,
since we are solving constant-coefficient linear elliptic partial
differential equations (PDEs), the solutions can be represented with
layer potentials
\begin{align}
  \label{eq:LP}
  u(\mathbf{x}) = \mathcal{K}[\sigma](\mathbf{x}) = 
  \int_{\partial\Omega} K(\mathbf{x},\mathbf{y})
  \sigma(\mathbf{y})\,dS,
\end{align}
where $\sigma$ is an unknown density function, and
$K(\mathbf{x},\mathbf{y})$ is a derivative of the fundamental solution
of the PDE. By matching the prescribed boundary condition at
$\mathbf{x}_0 \in \partial\Omega$ with the limit of
equation~\eqref{eq:LP} as $\mathbf{x}\rightarrow \mathbf{x}_0$, a
boundary integral equation (BIE) for the density function is formed.

@@

Significant challenges to solving BIEs include solving largely stem from
solving dense linear systems. \S\ref{subsec:NumericalIssues}
describes how we will combine preconditioning strategies with fast
summation methods to accelerate the time to solution,
\S\ref{subsec:timeStepping} describes adaptive time stepping methods,
and \S\ref{subsec:AC} describes how the non-linear Allen-Cahn
equation~\eqref{eqn:???} will be iteratively solved using BIE
formulations.

@@

\subsubsection{Iterative methods for the Allen-Cahn HAP model}
\label{subsec:AC}
\todo[inline]{Write once the model is finalized}

>>>>

\subsection{Specific Aim 2: Efficient, high-order methods for
large-scale simulations}
% ----------------------------------------------------------------------
Over the past decade, there has been an explosion of interest in
small-scale processes that utilize capillary forces and van der Waals
interactions to coordinate movement and bind microscopic components in
solvent~\cite{Pandey2011, Zhang2017, Siontorou2017}. Additionally,
hydrodynamic interactions are central to fabricate complex,
three-dimensional microstructures~\cite{Dasgupta2017, Leong2007,
Reynolds2019, Cho2010}. Hydrodynamic effects cannot be ignored since
they, like for the rates of biological functions like pore
dynamics~\cite{RYHAM20112929}, set the rate of dissipation. In Specific
Aim 2, we propose to develop efficient high-order numerical methods to
simulate dynamic assembly of particles under HAP potential in a viscous
solvent. As described in~\eqref{eq:stokes1}--\eqref{eq:stokes3} and
illustrated in Figure~\ref{fig:flow_map}, the solvent phase is modeled
by the Stokes equations for an incompressible, viscous fluid, and these
equations are coupled to the screened-Laplace equation~\eqref{eq:SL}
through viscous and hydrophobic stress balance.

\label{subsec:specific_aim_2}
\begin{wrapfigure}[14]{r}{0.61\textwidth}
  \centering{\includegraphics[width=0.60\textwidth]{Figures/domain.pdf}}
    \vspace{-15pt}
    \caption{\label{fig:flow_map} \footnotesize A collection of 3D
    amphiphilic particles suspended in a solvent. The solvent velocity
    satisfies the Stokes equations, and the hydrophobic forces depend on
    the solvent activity that satisfies the screened Laplace equation.
    The geometry is updated by solving the mobility problem.} 
\end{wrapfigure}
The HAP model requires solving exterior Dirichlet problems for the order
parameter~\eqref{eq:SL} and the Stokes
equations~\eqref{eq:stokes1}--\eqref{eq:stokes2}. If discretizing these
equations with stencil-based numerical methods, the computational domain
must be truncated, the fluid volume must be discretized, and artificial
boundary conditions must be imposed. The artificial boundary conditions
introduce additional error, and discretizing the volume results in an
excessively large system of equations. Moreover, it is difficult to
obtain high-order discretizations when the boundary surfaces are
irregular.

Alternatively, we will take advantage of the linearity of the Stokes
equations, and also apply a Newton's method to the HAP
model~\eqref{eq:SL}, where each iteration requires the solution of
linear elliptic PDEs. Then we can represent solutions of these elliptic
PDEs as layer potentials
\begin{align}
  \label{eq:LP}
  u(\mathbf{x}) = \mathcal{K}[\sigma](\mathbf{x}) = 
  \int_{\partial\Omega} K(\mathbf{x},\mathbf{y})
  \sigma(\mathbf{y})\,dS,
\end{align}
where $\sigma$ is an unknown density function, and
$K(\mathbf{x},\mathbf{y})$ is a derivative of the fundamental solution
of the PDE. By matching the prescribed boundary condition at
$\mathbf{x}_0 \in \partial\Omega$ with the limit of
equation~\eqref{eq:LP} as $\xx \rightarrow \xx_0$, a boundary integral
equation (BIE) for the density function is formed.

@@

Significant challenges to solving BIEs include solving largely stem from
solving dense linear systems. \S\ref{subsec:NumericalIssues}
describes how we will combine preconditioning strategies with fast
summation methods to accelerate the time to solution,
\S\ref{subsec:timeStepping} describes adaptive time stepping methods,
and \S\ref{subsec:AC} describes how the non-linear Allen-Cahn
equation~\eqref{eq:SL} will be iteratively solved using BIE
formulations.

\subsubsection{Iterative methods for the Allen-Cahn HAP model}
\label{subsec:AC}
In the PIs collaboration, they have developed numerical methods to
solve~\eqref{eq:SL} when $f(u) = u^2$. With this choice of $f(u)$, the
resulting PDE is the linear screened Laplace equation which is
efficiently solved with a double-layer potential representation. For a
more general $f$, such as a double well potential, we will interpret the
solution of~\eqref{eq:SL} as a root finding problem for the operator
\begin{align*}
  F[u] = -\rho \Delta u + f'(u).
\end{align*}

@@
